---
title: "brazil_legacy_cleaning"
author: "Jake Eisaguirre"
date: "2022-08-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

# librarian downloads, if not already downloaded, and reads in needed packages

librarian::shelf(here, tidyverse, janitor, lubridate, parsedate, stringr,hms, stringi, readr)
```

# load brazil data and rename columns
```{r}

brazil <- read_csv(here("data", "brazil_bd.csv")) %>% 
  clean_names() %>% 
  rename(date = date_formatted,
         survey_time = day_night,
         area = transect_id,
         environment = stream_distance,
         swab_id = swab,
         comments = observations_portuguese) %>% 
  mutate(area = str_replace(area, "S", ""))

```

# load campaign data
```{r}

camp <- read_csv(here("data", "brazil_sites.csv")) %>% 
  mutate(latitude = as.character(str_sub(latitude, end = 9)),
         longitude = as.character(str_sub(longitude, end = 9)),
         latitude = -1 * as.numeric(latitude),
         longitude = -1 * as.numeric(longitude),
         area = str_to_lower(area)) %>% 
  pivot_wider(names_from = start_end, values_from = c(latitude, longitude))
  

write_csv(camp, here("clean_tables", "campaign_location.csv"))
  
```

# clean dates per campaign. Different date formats.... lame
```{r}

camp_1 <- brazil %>% 
  filter(campaign == "1") %>% 
  mutate(date = lubridate::parse_date_time(date, c("%d/%m/%y"))) %>% 
  select(!date_unformatted)

camp_2 <- brazil %>% 
  filter(campaign == "2") %>% 
  mutate(date = lubridate::parse_date_time(date, c("%d/%m/%y")))%>% 
  select(!date_unformatted)

camp_3 <- brazil %>% 
  filter(campaign == "3") %>% 
  mutate(date = lubridate::parse_date_time(date, c("%d/%m/%y")))%>% 
  select(!date_unformatted)

camp_4 <- brazil %>% 
  filter(campaign == "4") %>% 
  mutate(date = lubridate::parse_date_time(date, c("%d/%m/%y"))) %>% 
  select(!date_unformatted)

camp_5 <- brazil %>% 
  filter(campaign == "5") %>% 
  mutate(date = str_replace(date, "/19/", "/9/"),
    date = lubridate::parse_date_time(date, c("%d/%m/%y")))%>% 
  select(!date_unformatted)

camp_6 <- brazil %>% 
  filter(campaign == "6") %>% 
  mutate(date = lubridate::parse_date_time(date, c("%d/%m/%y"))) %>% 
  select(!date_unformatted)

camp_7 <- brazil %>% 
  filter(campaign == "7") %>% 
  mutate(date = lubridate::parse_date_time(date, c("%d/%m/%y")))%>% 
  select(!date_unformatted)

camp_8 <- brazil %>% 
  filter(campaign == "8") %>% 
  mutate(date = lubridate::parse_date_time(date, c("%d/%m/%y")))%>% 
  select(!date_unformatted)

camp_9 <- brazil %>% 
  filter(campaign == "9") %>% 
  mutate(date = lubridate::parse_date_time(date, c("%d/%m/%y")))%>% 
  select(!date_unformatted)

camp_10 <- brazil %>% 
  filter(campaign == "10") %>% 
  mutate(date = lubridate::parse_date_time(date, c("%d/%m/%y")))%>% 
  select(!date_unformatted)

camp_11 <- brazil %>% 
  filter(campaign == "11") %>% 
  mutate(date = lubridate::parse_date_time(date, c("%d/%m/%y")))%>% 
  select(!date_unformatted)


brazil <- rbind(camp_1, camp_2, camp_3, camp_4, camp_5, camp_6, camp_7, camp_8, camp_9, camp_10, camp_11)
  
```


# convert characters to english and fix lat/long
```{r}

brazil_english <- brazil %>% 
  mutate(survey_time = str_to_lower(survey_time),
         environment = str_to_lower(environment),
         family = str_to_lower(family),
         species = str_to_lower(species),
         area = str_to_lower(area),
         latitude = as.character(str_sub(latitude, end = 9)),
         latitude = abs(as.numeric(latitude)),
         latitude = (-1 *latitude),
         longitude = as.character(str_sub(longitude, end = 9)),
         longitude = abs(as.numeric(longitude)),
         longitude = (-1 *longitude))

```

# split apart start_time column
```{r}

b_start_end <- brazil_english %>% 
  mutate(end_time = str_sub(time_beginning_end, -5),
         start_time= str_sub(time_beginning_end, start = 1, end = -7),
         count = as.numeric(1)) %>% 
  relocate(start_time, .before = capture_time) %>% 
  relocate(end_time, .after = start_time) %>%
  relocate(count, .after = capture_time) %>% 
  select(!time_beginning_end) %>%
  mutate(end_time = str_replace(end_time, "/", "0"))
  
  

```

# convert columns to bool
```{r}


brazil_bool <- b_start_end %>% 
  mutate(glycerol_swab_sample = if_else(glycerol_swab_sample == "x", "1", "0", missing = "0"),
         tissue_sample = if_else(tissue_sample == "x", "1", "0", missing = "0"),
         marked_toeclip = if_else(marked_toeclip == "x", "1", "0", missing = "0")) %>% 
  rename(capture_latitude = latitude,
         capture_longitude = longitude)


```

# fix end_time leading zero issue and add seconds zeros
```{r}
brazzzzzil<- data.frame(gsub("\\b(\\d{1})\\b", "0\\1", as.matrix(brazil_bool)))

brazil_time_fix <- brazzzzzil %>% 
  mutate(start_time = if_else(start_time == "", NA, start_time),
         start_time = paste(start_time, ":00", sep = ""),
         end_time = paste(end_time, ":00", sep = ""),
         #capture_time = paste(capture_time, ":00", sep = ""),
         start_time = str_replace(start_time, "NA:00", "NA"),
         end_time = str_replace(end_time, "NA:00", "NA"),
         count = as.numeric(count),
         glycerol_swab_sample = as.numeric(glycerol_swab_sample),
         tissue_sample = as.numeric(tissue_sample),
         marked_toeclip = as.numeric(marked_toeclip),
         location = "brazil",
         date = parse_date(date),
         area = str_replace(area, "0", ""),
         campaign = as.numeric(campaign)) %>% 
  left_join(camp, by = c("area", "environment")) %>% 
  rename(glycerol = glycerol_swab_sample,
         tissue = tissue_sample,
         marked = marked_toeclip,
         scientific_name = species,
         time_of_capture = capture_time)

```

# fix mis-inputted end_time
```{r}

# b <- brazil_time_fix %>% 
#   select(c(start_time, end_time, time_of_capture)) %>% 
#   mutate(start_time = sort(start_time))
# 
# row_sap <- brazil_time_fix %>%
#   mutate(mint = pmax(time_of_capture, end_time),
#          end_time = pmax(start_time, end_time),
#          start_time = mint, mint = NULL)
# 
# brazil_time_fix <- left_join(brazil_time_fix, row_sap, by = "swab")
# 
# 
# 
# duration_brazil <- brazil_time_fix %>% 
#   mutate(start_hour = as_hms(start_time),
#          end_hour = hour(end_time),
#          duration_min = if_else(end_time < start_time,
#                             as_hms(86400) - start_time + end_time,
#                             end_time - start_time),
#          duration_min = duration_min/60)
# 
# a <- brazil_time_fix %>% 
#   mutate(start_time = as_hms(start_time))

```

# location table
```{r}
location_table <- brazil_time_fix %>% 
  select(location) %>% 
  group_by(location) %>% 
  mutate(loc_temp_id = cur_group_id()) %>% 
  filter(!duplicated(loc_temp_id))

write_csv(location_table, here("clean_tables", "location.csv"))
```


# create campaign table
```{r}

campaign_table <- brazil_time_fix %>% 
  select(campaign, location) %>% 
  group_by(campaign) %>% 
  mutate(camp_temp_id = cur_group_id()) %>% 
  filter(!duplicated(camp_temp_id))

write_csv(campaign_table, here("clean_tables", "campaign.csv"))

```

# create site table
```{r}

site_table <- brazil_time_fix %>% 
  select(area, environment, longitude_start, 
         longitude_end, latitude_start, latitude_end) %>% 
  group_by(area, environment) %>% 
  mutate(site_temp_id = cur_group_id()) %>% 
  filter(!duplicated(site_temp_id)) %>% 
  select(!c(site_temp_id))

write_csv(site_table, here("clean_tables", "site.csv"))

```


# create visit table
```{r}

visit_table <- brazil_time_fix %>% 
  select(campaign, date, area, survey_time, environment) %>% 
  unite(site, c(area, environment), sep="_") %>% 
  group_by(date, survey_time, site) %>% 
  mutate(v_temp_id = cur_group_id()) %>% 
  filter(!duplicated(v_temp_id)) %>% 
  mutate(site = str_replace(site, " ", "_")) %>% 
  select(!c(v_temp_id))

write_csv(visit_table, here("clean_tables", "visit.csv"))

```

# create bad survey table but following structure
```{r}

survey_table <- brazil_time_fix %>% 
  select(start_time, end_time, environment, date, survey_time, area, campaign) %>% 
  unite(site, c(area, environment), sep="_")  %>%
  group_by(date, site, survey_time) %>% 
  mutate(temp_surv_id = cur_group_id()) %>% 
  filter(!duplicated(temp_surv_id)) %>% 
  mutate(site = str_replace(site, " ", "_"),
         start_time = if_else(start_time == "NA", NA, start_time),
         end_time = if_else(start_time == "NA", NA, end_time)) %>% 
  select(!c(temp_surv_id))

write_csv(survey_table, here("clean_tables", "survey.csv"))

```

# capture table
```{r}

capture_table <- brazil_time_fix %>% 
  select(!c(location, start_time, end_time)) %>% 
  mutate(glycerol = as.logical(glycerol),
         tissue = as.logical(tissue),
         marked = as.logical(marked),
         sex = if_else(sex %in% c("FO", "F"), "female", "male")) %>% 
  select(!c(dead, c(bd_presence:log10_1)))

write_csv(capture_table, here("clean_tables", "capture.csv"))

```

# pull out new bd and CMR data
```{r}

cmr <- capture_table %>% 
  select(swab_id, sex, recapture_from_campaign_n_number) %>% 
  rename(capture_mark_recapture = recapture_from_campaign_n_number) %>% 
  write_csv(here("clean_tables", "cmr.csv"))

# bd <- capture_table %>% 
#   select(swab_id, zoo_load_100, log10_1) %>% 
#   write_csv(here("clean_tables", "bd.csv"))

```

