---
title: "brazil_legacy_cleaning"
author: "Jake Eisaguirre"
date: "2022-08-22"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

if (!require(librarian)){
  install.packages("librarian")
  library(librarian)
}

# librarian downloads, if not already downloaded, and reads in needed packages

librarian::shelf(here, tidyverse, janitor, lubridate, parsedate, stringr,hms, stringi)
```

# load brazil data and rename columns
```{r}

brazil <- read_csv(here("data", "brazil_bd.csv")) %>% 
  clean_names() %>% 
  rename(campaign = campanha,
         date = data,
         survey_time = dia_noite,
         area = area_amostral,
         environment = ponto,
         family = familia,
         scientific_name = nome_cientifico,
         time_of_capture = horario,
         swab_id = swab,
         start_end = inicio_fim,
         glycerol = glicerol,
         tissue = tecido,
         marked = marcado,
         comments = observacoes) %>% 
  mutate(area = str_replace(area, "S", ""))

```

# load campaign data
```{r}

camp <- read_csv(here("data", "brazil_sites.csv")) %>% 
  mutate(latitude = as.character(str_sub(latitude, end = 9)),
         longitude = as.character(str_sub(longitude, end = 9)),
         latitude = -1 * as.numeric(latitude),
         longitude = -1 * as.numeric(longitude),
         area = str_to_lower(area)) %>% 
  pivot_wider(names_from = start_end, values_from = c(latitude, longitude))
  

write_csv(camp, here("clean_tables", "campaign_location.csv"))
  
```


# convert characters to english and fix lat/long
```{r}

brazil_english <- brazil %>% 
  mutate(survey_time = ifelse(survey_time=="Dia", "day", "night"),
         environment = ifelse(environment=="Terra", "land", "water"),
         date = parse_date(date),
         family = str_to_lower(family),
         scientific_name = str_to_lower(scientific_name),
         area = str_to_lower(area),
         latitude = as.character(str_sub(latitude, end = 9)),
         latitude = abs(as.numeric(latitude)),
         latitude = (-1 *latitude),
         longitude = as.character(str_sub(longitude, end = 9)),
         longitude = abs(as.numeric(longitude)),
         longitude = (-1 *longitude))

```

# split apart start_time column
```{r}

b_start_end <- brazil_english %>% 
  mutate(end_time = str_sub(start_end, -5),
         start_time= str_sub(start_end, start = 1, end = -7),
         count = as.numeric(1)) %>% 
  relocate(start_time, .before = time_of_capture) %>% 
  relocate(end_time, .after = start_time) %>%
  relocate(count, .after = time_of_capture) %>% 
  select(!start_end) %>%
  mutate(end_time = str_replace(end_time, "/", "0"))
  
  

```

# convert columns to bool
```{r}


brazil_bool <- b_start_end %>% 
  mutate(glycerol = if_else(glycerol == "x", "1", "0", missing = "0"),
         tissue = if_else(tissue == "x", "1", "0", missing = "0"),
         marked = if_else(marked == "x", "1", "0", missing = "0")) %>% 
  rename(capture_latitude = latitude,
         capture_longitude = longitude)


```

# fix end_time leading zero issue and add seconds zeros
```{r}
brazzzzzil<- data.frame(gsub("\\b(\\d{1})\\b", "0\\1", as.matrix(brazil_bool)))

brazil_time_fix <- brazzzzzil %>% 
  mutate(start_time = paste(start_time, ":00", sep = ""),
         end_time = paste(end_time, ":00", sep = ""),
         time_of_capture = paste(time_of_capture, ":00", sep = ""),
         start_time = str_replace(start_time, "NA:00", "NA"),
         end_time = str_replace(end_time, "NA:00", "NA"),
         count = as.numeric(count),
         glycerol = as.numeric(glycerol),
         tissue = as.numeric(tissue),
         marked = as.numeric(tissue),
         location = "brazil",
         date = parse_date(date),
         area = str_replace(area, "0", "")) %>% 
  left_join(camp, by = c("area", "environment"))

```

# fix mis-inputted end_time
```{r}

# b <- brazil_time_fix %>% 
#   select(c(start_time, end_time, time_of_capture)) %>% 
#   mutate(start_time = sort(start_time))
# 
# row_sap <- brazil_time_fix %>%
#   mutate(mint = pmax(time_of_capture, end_time),
#          end_time = pmax(start_time, end_time),
#          start_time = mint, mint = NULL)
# 
# brazil_time_fix <- left_join(brazil_time_fix, row_sap, by = "swab")
# 
# 
# 
# duration_brazil <- brazil_time_fix %>% 
#   mutate(start_hour = as_hms(start_time),
#          end_hour = hour(end_time),
#          duration_min = if_else(end_time < start_time,
#                             as_hms(86400) - start_time + end_time,
#                             end_time - start_time),
#          duration_min = duration_min/60)
# 
# a <- brazil_time_fix %>% 
#   mutate(start_time = as_hms(start_time))

```

# location table
```{r}
location_table <- brazil_time_fix %>% 
  select(location) %>% 
  group_by(location) %>% 
  mutate(loc_temp_id = cur_group_id()) %>% 
  filter(!duplicated(loc_temp_id))

write_csv(location_table, here("clean_tables", "location.csv"))
```


# create campaign table
```{r}

campaign_table <- brazil_time_fix %>% 
  select(campaign, location) %>% 
  group_by(campaign) %>% 
  mutate(camp_temp_id = cur_group_id()) %>% 
  filter(!duplicated(camp_temp_id))

write_csv(campaign_table, here("clean_tables", "campaign.csv"))

```

# create site table
```{r}

site_table <- brazil_time_fix %>% 
  select(campaign, location, date, area, environment, survey_time, longitude_start, 
         longitude_end, latitude_start, latitude_end) %>% 
  group_by(area, environment) %>% 
  mutate(site_temp_id = cur_group_id()) %>% 
  filter(!duplicated(site_temp_id))

write_csv(site_table, here("clean_tables", "site.csv"))

```


# create visit table
```{r}

visit_table <- brazil_time_fix %>% 
  select(campaign, date, area, survey_time, environment) %>% 
  group_by(date, survey_time) %>% 
  mutate(v_temp_id = cur_group_id()) %>% 
  filter(!duplicated(v_temp_id))

write_csv(visit_table, here("clean_tables", "visit.csv"))

```

# create bad survey table but following structure
```{r}

survey_table <- brazil_time_fix %>% 
  select(start_time, end_time, environment, date, survey_time, area, campaign) %>% 
  group_by(date, area, survey_time, environment) %>% 
  mutate(temp_surv_id = cur_group_id()) %>% 
  filter(!duplicated(temp_surv_id))

write_csv(survey_table, here("clean_tables", "survey.csv"))

```

# capture table
```{r}

capture_table <- brazil_time_fix %>% 
  select(!c(location, start_time, end_time)) %>% 
  mutate(glycerol = as.logical(glycerol),
         tissue = as.logical(tissue),
         marked = as.logical(marked))

write_csv(capture_table, here("clean_tables", "capture.csv"))

```

